 
 \documentclass{article}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{fancyhdr}
\usepackage{fasy-hw}

\pagestyle{fancy}
\fancyhf{}
\lhead{Project 3: Burrows-Wheeler Transform}
\rhead{Troy Oster, George Engel, Gavin Austin, Thomas Herndon}

\begin{document}
   The Burrows-Wheeler transform is an algorithm created by Michael Burrows and David Wheeler in 1994 at the DEC Systems Research Center in California. It is primarily used in compression, as it transforms a character string into other strings of similar characters.[1] This algorithm was designed to solve a specific problem with compression, that is, the costly nature of data preparation for use in the data compression algorithms/techniques at the time. The Burrows-Wheeler transform algorithm solves this problem quite elegantly, simply by being faster than every other known data preparation technique when it was created.[1]

   The important aspect to note about the transformation produced by this algorithm is that it is reversible. To be fully reversed, the transformation only needs to store the position of the original first character.[2] The Burrows-Wheeler transform improves the efficiency of almost all text compression algorithms, as it only costs a small amount of extra computation.[2] Due to this ability to improve compression efficiency, it is often used to prepare the data used in data/text compression techniques. 
   
   The basic explanation of the Burrows-Wheeler algorithm is that it takes in a string as an input and outputs a string composed of similar letters to those present in the input string. The output string is then composed of many repeated characters, which makes it easier to compress than the original input. This, through an optimization of the input, optimizes (almost) any compression algorithm used.[2]
   
   The transform performed by the Burrows-Wheeler algorithm is completed by creating circular shifts of text and then sorting them by lexicographic order. After sorting is complete, the last column and the index of the input string in the set of sorted permutations are extracted. The characters found in the last column are then concatenated and the resulting string is the output of the Burrows-Wheeler transform. [2]
   
   This algorithm, despite having been invented in 1994, is still used in many fields of research today. The primary field of use for the Burrows-Wheeler algorithm is Bioinformatics. Variations of this algorithm and more complex implementations of it, such as the Burrows-Wheeler Alignment (BWA), have been shown to be roughly ten to twenty times faster than other commonly used methods for alignment.[3]
   
   The pseudo-code for the implementation of the Burrows-Wheeler algorithm is shown below.[2]
   
\textbf{Pseudo-Code}

    \begin{algorithm}
    \caption{Reversible Transformation}
    \textbf{Input:} A string $S$ of $N$ characters $\in$ some character alphabet $X$
    
    \textbf{Output:} A column $L$ of the character matrix
        \begin{algorithmic}[1]
            \Procedure{Compression\_Transform}{$S$}
            \State M $\gets$ char[N][N]      \Comment{Define M as an N x N Matrix of chars}
            \For{int i = 0; i $<$ len(S); i++}
                \State M[i]$ \gets$ Shift(S,i) \Comment{Sets M[i] to the char set shifted by i}
            \EndFor
            \State lexicographic\_sort(M)
            
            \State $L \gets M.col(-1)$ \Comment{Set L as the last column from the matrix}
            \State \textbf{return} $L$
            \EndProcedure
        \end{algorithmic}
    \end{algorithm}
\newpage
    \begin{algorithm}
        \caption{Decompression Transformation}
        \textbf{Input:} The output column $L$ from Algorithm 1
    
    \textbf{Output:} A string $S$
        \begin{algorithmic}[1]
            \Procedure{Decompression\_Transform}{$L$}
            \State $F \gets lexicographic\_sort(L)$
            \State initialize($M$)\Comment{A table that will eventually hold M from Algorthim 1}
            \For{int i = 0; i $<$ len(L); i++}
                \State M.prepend(S) \Comment{Prepend the string $L$ to the table $M$}
                \State lexicographic\_sort(M)
            \EndFor
           \State $I\gets 0$
            \While{M[I][n-1]!=EOF}
                \State $I\gets I+1$
            \EndWhile
            \State $S \gets M[I]$ \Comment{Set S as the row that ends in a end of file character}
            \State $return$ $S$
            \EndProcedure
        \end{algorithmic}
    \end{algorithm}
       
   \section{References}
   \makebox[.5cm]{[1]} Adjeroh, Donald, et al. The Burrows-Wheeler Transform: Data Compression, Suffix Arrays, and Pattern Matching. Springer, 2011.\par
    \makebox[.5cm]{[2]}  Burrows, Michael; Wheeler, David J. (1994), A block sorting lossless data compression algorithm, Technical Report 124, Digital Equipment Corporation\par
    \makebox[.5cm]{[3]} Heng Li, Richard Durbin, Fast and accurate short read alignment with Burrows–Wheeler transform, Bioinformatics, Volume 25, Issue 14, 15 July 2009, Pages 1754–1760, https://doi.org/10.1093/bioinformatics/btp324\par
\end{document}